### 2018-04-29 paper/evaluation.tex

Each round lasts $T=10s$ \commentAL{yes, the user switches places fast} and users start at different times. 
This way, even though the round's period is fixed, devices enter and leave the system at different points in the continuous time.

@@

>>>>

Each round lasts $T=10s$ \commentAL{yes, the user switches places fast} and users start at different times. 
This way, even though the round period is fixed, devices enter and leave the system at different points in the continuous time.

@@

\subsection{Devices' predictions of the user behavior} % (fold)
\label{sub:devices_predictions_of_the_user_behavior}

% subsection devices_predictions_of_the_user_behavior (end)
\commentAL{Should evaluate the inference of the model given only the observable sequence. Also show that $P[O_{t+1} | S_t] \simeq P[O_{t+i} | S_t]$}

### 2018-04-29 paper/squad-overlay-design.tex

To this end, we model the user's behavior as a Hidden Markov Models (HMMs) of order one, with multiple discrete \emph{observable} processes (one per device) stored in $O$. 
The hidden process $S=S_1, ..., S_i, ...$ represents the user's locations, constituted of a set of $N$ potential locations (e.g. home, work or outside), $N$ being a user-defined parameter.

@@

\begin{figure}[t]
\centering
\vspace{-1em}

$$A =
\kbordermatrix{
      & W            & O            & H            \cr
    W & \sfrac{2}{3} & \sfrac{1}{3} & 0            \cr
    O & \sfrac{1}{3} & \sfrac{1}{3} & \sfrac{1}{3} \cr
    H & 0            & \sfrac{1}{3} & \sfrac{2}{3} \\[0.3em]
}, \;
B = 
\kbordermatrix{
      & W     & O   & H   \cr
    o_p & 0.7 & 0.6 & 0.8 \cr
    o_w & 0.7 & 0   & 0   \cr
    o_h & 0   & 0   & 0.7 \cr
    o_l & 0.2 & 0.4 & 0.6 \\[0.3em]
}$$

\includegraphics[width=0.9\columnwidth]{figures/hmm.pdf}
\caption{ \label{fig:hmm} Example Hidden Markov Model (HMM) of a user's behavior.}
\end{figure}

We consider a single user, owning a set of $\mathcal{D}$ devices.
Time is discretized \commentAL{Discretized/synchronous time needs more explanation}. Needs, and the devices' observations are independent: Alice can have her laptop and phone connected at the same time.
The hidden process $S=S_1,\dots,S_T$ is built from an alphabet of states $\mathcal{S}$, while the observable process $O=O_1, \dots, \O_T$ contains events from the alphabet $\mathcal{O}$. We define a multivariate HMM $\lambda_N$ as follows:

@@

	\item $B = \left\{ b_{i\,k}\right\}_{(i,k)\in\mathcal{S}\times\mathcal{O}}$ is the matrix of event emission probabilities for each state:
	$b_{i\,k} = P[ k \in O_t \mid S_t = i]$.
\end{itemize}

In our case, the states are the different locations where the user is susceptible to go;
we considered the three following ones: $\mathcal{S}=\left\{ \mathit{work}, \mathit{outside}, \mathit{home} \right\}$. 
There is one event per device $d$: $\mathcal{O} = \left\{ o_d \right\}_{d\in \mathcal{D}}$. Several devices can be connected at the same time, such that $O_t$ is a set, and $o_d \in O_t$ means that device $d$ was connected at time $t$. Devices' connection probabilities are independent.

In \cref{fig:hmm}, we show an example HMM of a user's behavioral HMM, along with its graphical representation (the initial state distribution $\Pi$ was left out for simplicity). 
The represented user possesses a phone and a laptop, that she carries around with her, a home computer, that is only accessible from her home, and a workstation, located at her workplace. She never forgets to turn off her fixed appliances before leaving.

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{figures/sample_usage.pdf}

\caption{\label{fig:sample_usage}Sample device usages using the previous HMM specifications. 
%The bottom line reads the HMM's states: \textbf{W}ork, \textbf{O}utside, and \textbf{H}ome, while the dark lines read when the devices (\textbf{h}ome computer, \textbf{p}hone, and \textbf{w}orkstation) are online.
}

\end{figure}

% Using the toggle events, we compute the device's \emph{connection} events, represented by the multivariate Bernoulli process $C=\left\{ c^{(d)} \right\}_{d\in\mathcal{D}}$. 
% $c^{(d)}_t=1$ (resp. 0) means that the device $d$ was online (resp. offline) at time step $t$.

% We state that every device starts offline, and compute $c^{(d)}_t$ as follows:

% $$ \forall d \in \mathcal{D},
% \begin{cases}
% c^{(d)}_t=0 &\text{when } t=0; \\
% c^{(d)}_t=c^{(d)}_{t-1} \oplus o^{(d)}_t &\text{when } t>0.
% \end{cases}$$

Figure~\ref{fig:sample_usage}, shows a possible trace of the user's behavior, using the parameters depicted in figure~\ref{fig:hmm}. 
It reads the timeline of $o_{d}$ for each device $d$. On the bottom of the graph, the first letter of the user's location is shown at each time step.
We see that devices connections depend on the user's location, and that any number of devices can be connected at the same time.

% Each device starts offline; then, the user either toggles one device per round or does nothing.
% In this figure, we show the device \emph{usage}, that is a consequence of the $\mathit{toggle}$ events defined by the HMM.
% As a consequence, several devices can be online at the same time, as happens in the figure.

% subsubsection formal_definition (end)

\subsubsection{Predicting future behavior} % (fold)
\label{ssub:predicting_future_behavior}
Devices want to predict whether they are likely to be connected in the near future.
This information will be publicly available, and will allow routes selection for the  files exchange between users.

@@

$$
P\left[ o_d \in O_{t} \mid S_{t-1} = s \right] = 
\sum\limits_{s' \in \mathcal{S}} 
a_{s\,s'} * b_{s'\,o_d}.
$$

@@

\begin{multline*}
P\left[ o_d \in O_{t+i} \mid S_{t} = s \right] = \\
\sum\limits_{s' \in \mathcal{S}}
P\left[ S_{t+1} = s' \mid S_t = s \right] * 
P\left[ o_d \in O_{t+i} \mid S_{t+1} = s'\right].
\end{multline*}

\commentAL{In conclusion: we want more unicorns at the Olympic Games.}

\commentAL{I don't mention prediction yet since I am not sure we get good enough results.}

>>>>

To this end, we model the user's behavior as a Hidden Markov Models (HMMs) of order one, with multiple discrete \emph{observable} processes (one per device) stored in $O$. 
The hidden process $S=S_1, ..., S_i, ...$ represents the user's locations, constituted of a set of $N$ potential locations (e.g. home, work or outside), $N$ being a user-defined parameter.
A HMM of order one is such that the hidden variable at time $t$, $S_t$, only depends on the hidden variable at $t-1$, $S_{t-1}$.
The observations at time $t$, $O_t$, depend on the current hidden variable only, $S_t$.

We first formally define our HMM, then show how to predict the future knowing the HMM, then finally present how to infer the model from the observable sequence only.

@@

%We consider a single user, owning a set of $\mathcal{D}$ devices.
%Time is discretized \commentAL{Discretized/synchronous time needs more explanation}. Needs, and the devices' observations are independent: Alice can have her laptop and phone connected at the same time.
The hidden process $S=S_1,\dots,S_i, \dots$ is built from an alphabet of states $\mathcal{S}$, while the observable process $O=O_1, \dots, \O_i, \dots$ contains elements from the alphabet $\mathcal{O}$. We define a multivariate HMM $\lambda_N$ as follows:

@@

	\item $B = \left\{ b_{i\,k}\right\}_{(i,k)\in\mathcal{S}\times\mathcal{D}}$ is the matrix of event emission probabilities for each state:
	$b_{i\,k} = P[ k \in O_t \mid S_t = i]$.
\end{itemize}

In our case, the states $\mathcal{S}$ are the different locations $\mathcal{L}$ where the user is susceptible to go.
The events represent the devices connection, hence $\mathcal{O} = \mathcal{D}$, where $\mathcal{D}$ is the set of the user's devices.
Several devices can be connected at the same time, such that $O_t$ is a set, and $d \in O_t$ means that device $d$ was connected at time $t$. Devices' connection probabilities are independent.

% In \cref{fig:hmm}, we show an example HMM of a user's behavioral HMM, along with its graphical representation (the initial state distribution $\Pi$ was left out for simplicity). 
% The represented user possesses a phone and a laptop, that she carries around with her, a home computer, that is only accessible from her home, and a workstation, located at her workplace. She never forgets to turn off her fixed appliances before leaving.

% Figure~\ref{fig:sample_usage}, shows a possible trace of the user's behavior, using the parameters depicted in figure~\ref{fig:hmm}. 
% It reads the timeline of $o_{d}$ for each device $d$. On the bottom of the graph, the first letter of the user's location is shown at each time step.
% We see that devices connections depend on the user's location, and that any number of devices can be connected at the same time.

\paragraph*{Predicting future behavior} 
As already stated, devices want to predict whether they are likely to be connected in the near future.
This information will be publicly available, and will enable the intelligent building of routes.

@@

$$
P\left[ d \in O_{t} \mid S_{t-1} = s \right] = 
\sum\limits_{s' \in \mathcal{S}} 
a_{s\,s'} * b_{s'\,d}.
$$

@@

\begin{multline*}
P\left[ d \in O_{t+i} \mid S_{t} = s \right] = \\
\sum\limits_{s' \in \mathcal{S}}
P\left[ S_{t+1} = s' \mid S_t = s \right] * 
P\left[ d \in O_{t+i} \mid S_{t+1} = s'\right].
\end{multline*}

We show in section~\ref{sub:devices_predictions_of_the_user_behavior} that $P\left[ d \in O_{t+1} \mid S_{t} = s \right] \simeq P\left[ d \in O_{t+i} \mid S_{t} = s \right]$. 
For that reason, devices only compute the prediction of their connection at $t+1$, that is considered a good enough estimator of their state ``in the near future''.

\paragraph*{Inferring the model from the sequence of observations}
